---
title: "Crs BC association Report"
author: "Robert Frömel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
params:
  rmd: "CrsBCReport.Rmd"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: false
    highlight: pygments 
editor_options: 
  chunk_output_type: console
---

```{r}
Sys.time()
```

```{r, include=FALSE}

# set path to package library
.libPaths("/path/to/snakemake/env/MPRA_R_processing/lib/R/library")

# set global chunk settings
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

suppressPackageStartupMessages({
  require(plyr)
  require(dplyr)
  require(readr)
  require(seqinr)
  require(stringr)  
  require(ggplot2)
  require(DT)
  require(htmltools)
})

# get the directory where Snakefile is stored
directory <- gsub("/[^/]*$", "", snakemake@scriptdir)

# Load in wildcards
FastqDir <- snakemake@wildcards$FastqDir
Design <- snakemake@wildcards$Design
LibraryName <- snakemake@wildcards$LibraryName
OutDir <- snakemake@input$od

# directory where plots will be exported
plots_dir <- file.path(OutDir, FastqDir, "plots")

# create directory to store plots
dir.create(plots_dir, showWarnings = F)
```


# GRE-BC asociation of **`r snakemake@wildcards$FastqDir`**.


```{r ReadMapped}
mapped <-read_csv(snakemake@input$mapped)
```

<br>

We have **`r nrow(mapped)`** barcodes-CRS pairs

```{r PlotReads}
readhist <- ggplot(mapped) + geom_histogram(aes(x=log10(READS)), stat = "count") + ylab("Frequency") + xlab("Log10(Reads)") + theme_bw()
readhist
```

```{r }
# Save the plot
ggsave(readhist,
       filename = file.path(plots_dir, "Reads_histogram.png"),
       height = 8, width = 7)
```

<br>
How about alignment quality / synthesis errors in the CRS library? <br>
**`r sum(mapped$MEANMATCHES >= 290 & mapped$MEANMATCHES <= 292)`** elements have a quality between 290 adn 292  and <br> 
**`r sum(mapped$MEANMATCHES <= 250)`** have a quality below 250.

```{r SeqStats}
Matchhist <- ggplot(mapped) + geom_histogram(aes(x=MEANMATCHES)) + ylab("log10(Frequency)") + xlab("Mean read quality") + theme_bw() + scale_x_continuous(limits=c(250,300), oob = scales::squish) + scale_y_log10()
Matchhist
```

```{r }
# Save the plot
ggsave(Matchhist,
       filename = file.path(plots_dir, "Match_histogram.png"),
       height = 8, width = 7)
```

In terms of volume, **`r round(100*sum(mapped$READS[mapped$MEANMATCHES>290])/sum(mapped$READS),digits=2)`%** of the library has a maximum of 2 mismatches. 

```{r SubsetMapped}
filtered <- mapped
filtered <- subset(filtered, READS > DEVIANTREADS * 5 & CRS != "*") #apply filters: reads > deviant reads * 5, matches a CRS
filtered <- subset(filtered, MEANMATCHES >= 290)
filtered <- subset(filtered, MEANMATCHES <= 292)
```

**`r nrow(filtered)`** barcodes-CRS pairs pass the DEVIANTREADS and MEANMATCHES quality filter

```{r FilterHomoNucleotides}
filtered <- filtered %>%
  filter(!str_count(.$BARCODE, pattern = "A") >= 10) %>%
  filter(!str_count(.$BARCODE, pattern = "C") >= 10) %>%
  filter(!str_count(.$BARCODE, pattern = "G") >= 10) %>%
  filter(!str_count(.$BARCODE, pattern = "T") >= 10) 
```

After subsetting for Homonucleotides, we still have **`r nrow(filtered)`** barcodes-CRS pairs

```{r WriteOutputCSVFile}
write_csv(filtered, file = gzfile(paste0(OutDir,"/", FastqDir, "/", Design, "_", LibraryName, "_mapped_filtered.csv.gz", sep="", collapse="")))
```

# GRE statistics

```{r GREStatistics}
crs <- ddply(filtered, "CRS", summarise, n = length(BARCODE), reads = sum(READS))
```

```{r WriteOutputRDSFile}
saveRDS(crs, paste0(OutDir,"/", FastqDir, "/", Design, "_", LibraryName, "_crs.rds", sep="", collapse=""))
```

```{r ReadFasta}
fasta <- read.fasta(snakemake@input$fasta)
meta_info <- read_csv(snakemake@input$meta)
```

**`r nrow(crs)`** out of **`r length(fasta)`** CRS are represented in the library. <br>
This correspodns to: **`r nrow(crs)/length(fasta)`% ** <br>

How many barcodes:

<br>

```{r GREHistogram, fing.width=6, fig.height=4.5}
GREHistogram <- ggplot(crs) + geom_histogram(aes(x=n), stat = "count") + ylab("Frequency") + xlab("Number of Barcodes") + theme_bw()
GREHistogram
```

```{r }
# Save the plot
ggsave(GREHistogram,
       filename = file.path(plots_dir, "GRE_histogram.png"),
       height = 8, width = 7)
```

## BC mapping

```{r GREBCSummary}
print(summary(crs$n))
```

```{r GREQuantile}
crs$rep <- with(crs, reads / sum(reads))
crs$nrep <- with(crs, rep/quantile(rep, 0.05))
```

## Quantile plot

```{r GREQuantilePlot}
GREQuantilePlot <- qplot(x = 1:nrow(crs), y =nrep , data=crs[order(crs$rep),]) + xlab("Representation of the CRS in the library") + 
  geom_vline(xintercept = c(0.05, 0.95) * nrow(crs)) + geom_hline(yintercept = quantile(crs$nrep, 0.95)) + scale_y_log10(breaks = quantile(crs$nrep, c(0,0.05,0.95,1)), labels = round(quantile(crs$nrep, c(0,0.05,0.95,1)), digits=2))+ theme_bw()
GREQuantilePlot
```

```{r }
# Save the plot
ggsave(GREQuantilePlot,
       filename = file.path(plots_dir, "GRE_Quantile.png"),
       height = 8, width = 7)
```

<br>
And what about the representation:

```{r GRERepresentationHistogram}
GRERepresentationHist <- ggplot(crs) + geom_histogram(aes(x=rep), stat = "count") + ylab("Frequency") + xlab("Representation of the CRS in the library") + theme_bw()
GRERepresentationHist
```

```{r }
# Save the plot
ggsave(GRERepresentationHist,
       filename = file.path(plots_dir, "GRE_representation_histogram.png"),
       height = 8, width = 7)
```

## Read statistics

```{r GREQuantileSummary}
print(summary(crs$reads/sum(crs$reads)))
```

# Visualizing key design features for *`r FastqDir`**.

<br>

```{r }
crs_merge <- crs
colnames(crs_merge)[1] <- c("SeqID")
crs_merge <- inner_join(crs_merge, meta_info, by = "SeqID")
```

```{r}
Feature_tf_boxplot <- ggplot(crs_merge ) + 
        geom_boxplot(aes(x=Universal.FirstTF, y = reads / sum(reads))) + 
        ylab("Reads / sum(Reads)") + 
        xlab("TFs") + 
        theme_bw()  + 
        theme(axis.text.x = element_text(angle = 90))

Feature_spacing_boxplot <- ggplot(crs_merge ) + 
        geom_boxplot(aes(x=Universal.Spacing, y = reads / sum(reads), group = Universal.Spacing)) + 
        ylab("Reads / sum(Reads)") + 
        xlab("Binding sites affinity") + 
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 90))

Feature_sublibrary_boxplot <- ggplot(crs_merge ) + 
        geom_boxplot(aes(x=SubLibrary, y = reads / sum(reads))) + 
        ylab("Reads / sum(Reads)") + 
        xlab("Sub Library") + 
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 90))
```

```{r}
Feature_tf_boxplot
```

```{r }
# Save the plot
ggsave(Feature_tf_boxplot,
       filename = file.path(plots_dir, "Feature_tf_boxplot.png"),
       height = 8, width = 7)
```

```{r}
Feature_spacing_boxplot
```

```{r }
# Save the plot
ggsave(Feature_spacing_boxplot,
       filename = file.path(plots_dir, "Feature_spacing_boxplot.png"),
       height = 8, width = 7)
```

```{r}
Feature_sublibrary_boxplot
```

```{r }
# Save the plot
ggsave(Feature_sublibrary_boxplot,
       filename = file.path(plots_dir, "Feature_sublibrary_boxplot.png"),
       height = 8, width = 7)
```

# Final Sumamry

**90%** of the library are represented with at least  **`r round(100*quantile(crs$reads/sum(crs$reads),0.1),digits=5)` %** , <br> 
that is, **`r round(100*quantile(crs$reads/sum(crs$reads),0.1)/max(crs$reads/sum(crs$reads)), digits=2)` %** of the most abundantly represented CRS <br> 
and **`r round(100*quantile(crs$reads/sum(crs$reads),0.1)/(1/10000), digits=2)`%** of what it would be in perfect world.


# Source
* <a download="CrsBCReport.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd',
    encoding = 'base64')`">R Markdown source file (to produce this document)</a>

