---
title: "CRS Barcode mapping"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic statistics


```{r read_data, fig.width=6, fig.height=4.5}
.libPaths(c("/home/lars/RLibs/R4.2.1/"))
require(plyr)
require(ggplot2)
mapped <- read.csv(gzfile("/users/lvelten/lvelten/Analysis/SCG4SYN/LibH/BCass/Analysis/mapped_filter.csv.gz"))
#mapped <- read.csv(gzfile("/users/lvelten/project/SCG4SYN/Experiments/221003_LibraryA_association/Analysis/merged_34_48/mapped_filter.csv.gz"))
#mapped <- read.csv(gzfile("/users/lvelten/project/SCG4SYN/Experiments/230308_LibraryH_association/Analysis/minP/mapped_filter.csv.gz"))
filtered <- na.omit(mapped)
```


```{r histo2, fig.width=6, fig.height=4.5}
nreads <- data.frame(reads = as.integer(names(table(filtered$READS))), instances = table(filtered$READS))
qplot(x = reads, y = instances.Freq, data = nreads, log = "xy") + xlab("Reads per barcodes") + ylab("Count")
```

There are `r nrow(filtered)` barcodes of which `r sum(filtered$READS >= 5) ` with at least 5 reads. In total `r sum(filtered$READS)` reads support a valid CRS-barcode match and `r sum(filtered$DEVIANTREADS)` reads support a secondary CRS-barcode match. This is only `r round(100*sum(filtered$DEVIANTREADS)/(sum(filtered$READS) + sum(filtered$DEVIANTREADS)), digits=1)` %.

Something is off here, because the libraries were sequenced to a depth of 280 Million? Looks like most barcodes (by far) were observed with one read. Count this to confirm in the pipeline.


```{r filter, fing.width=6, fig.height=4.5}
#apply filters
filtered <- mapped
 filtered <- subset(filtered, READS > 5)
filtered <- subset(filtered, READS > DEVIANTREADS * 5 & CRS != "*") #apply filters: reads > deviant rerads * 5, matches a CRS
```

How about alignment quality / synthesis errors in the CRS library?

```{r seqstat1, fing.width=6, fig.height=4.5}
qplot(x = MEANMATCHES, data = filtered, geom="histogram") + scale_x_continuous(limits=c(250,300), oob = scales::squish)
```

The vast majority of sequences does not have any mismatches. In terms of volume, `r round(100*sum(filtered$READS[filtered$MEANMATCHES>290])/sum(filtered$READS),digits=2)`% of the library has a maximum of 2 mismatches. We could check this against the synthesis quality that twist claims. What does QC_Failure mean in the design file / CRS name? There is no real effect of QC_Failure on the matches.

```{r filter2, fing.width=6, fig.height=4.5}
#apply filters
filtered <- subset(filtered, MEANMATCHES >= 290)
#write.csv(filtered, file = "outNextseq/mapped.filtered.csv",row.names=F,quote=F)
```

After this there are now `r nrow(filtered)` barcodes-CRS pairs

## CRS statistics

```{r crscount, fing.width=6, fig.height=4.5}
#remvoe the QC_Failure__ flag
filtered$CRS <- gsub("QC_Failure___", "", filtered$CRS)

crs <- ddply(filtered, "CRS", summarise, n = length(BARCODE), reads = sum(READS))
saveRDS(crs, file = "crs.rds" )
```

`r nrow(crs)` out of 10260 CRS are represented in the library. How many barcodes:

```{r crshist, fing.width=6, fig.height=4.5}
#apply filters
qplot(x = n, geom="histogram", data=crs) + xlab("Number of barcodes")

print(summary(crs$n))
```
The library is extremely complex, well done!

```{r crsreadhist, fing.width=6, fig.height=4.5}
#apply filters
crs$rep <- with(crs, reads / sum(reads))
crs$nrep <- with(crs, rep/quantile(rep, 0.05))
qplot(x = 1:nrow(crs), y =nrep , data=crs[order(crs$rep),]) + xlab("Representation of the CRS in the library") + 
  geom_vline(xintercept = c(0.05, 0.95) * nrow(crs)) + geom_hline(yintercept = quantile(crs$nrep, 0.95)) + scale_y_log10(breaks = quantile(crs$nrep, c(0,0.05,0.95,1)), labels = round(quantile(crs$nrep, c(0,0.05,0.95,1)), digits=2))
qplot(x = rep, geom="histogram", data=crs) + xlab("Representation of the CRS in the library")

print(summary(crs$reads/sum(crs$reads)))
```
Also the representation of the different CRS (in terms of volume) is not so even even: 90% of the library are represented with at least  `r round(100*quantile(crs$reads/sum(crs$reads),0.1),digits=5)` % , that is, `r round(100*quantile(crs$reads/sum(crs$reads),0.1)/max(crs$reads/sum(crs$reads)), digits=2)` % of the most abundantly represented CRS and `r round(100*quantile(crs$reads/sum(crs$reads),0.1)/(1/nrow(crs)), digits=2)`% of what it would be in perfect world.

```{r crsreadbox, fing.width=8, fig.height=4.5}
crs$round <- gsub(".+__","",crs$CRS)
crs$round[!crs$round %in% c("Round1","Round2","Round3","TForder_block")] <- "Controls"
qplot(x = round , y = reads / sum(reads), geom="boxplot", data=crs) + ylab("Representation of the CRS in the library") + theme(axis.text.x = element_text(angle = 90))
```

Analyze if there is a relationship between coverage and GC content

```{r crsreadseq, fig.width=8, fig.height=4.5}
require(stringr)
require(LSD)
Library1_renamed <- readRDS("/users/lvelten//lvelten/Analysis/SCG4SYN/LibH/221219_LibraryHeptad_for_ordering.rds")
crs$seq <- tolower(Library1_renamed[as.character(crs$CRS)])
crs$gc <- str_count(as.character(crs$seq), "[gc]") / nchar(crs$seq)
crs$gc.bin <- infotheo::discretize(crs$gc, nbins = 20)$X
heatscatter(x = crs$gc , y = crs$reads / sum(crs$reads))
qplot(x = as.factor(gc.bin) , y = reads / sum(reads), geom="boxplot", data=crs) + ylab("Representation of the CRS in the library")

```

Something wrong with the sequence names, get correct names from Robert